Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support).

### 2017-02-08 csf-2017/paper.tex
changed weak/strong to public/private in intro

In this paper, we make progress towards the goal of efficient and provably
watertight enforcement of encapsulation and control-flow integrity on capability
machines.  Specifically, we make the following contributions:
\begin{itemize}
\item We formalise a simple but representative capability machine featuring
  local capabilities much like CHERI's, and define its operational semantics.
\item We define a novel stack management discipline for enforcing control-flow
  integrity and encapsulation of local state on the stack. The discipline relies
  solely on local capabilities (used to represent stack and return pointers) and
  does not require OS support in the form of a trusted shadow stack or special
  call and return instructions. It also does not require separate per-component
  stacks and instead works with a single, fairly standard, contiguous stack. It
  supports higher-order cross-component calls (e.g., (potentially)
  cross-component function pointers) and can be efficient assuming
  only \lars{can we say only here ?}
  one additional piece of processor support: an instruction for efficiently clearing
  a range of memory.
\item We provide a theorem that is (by far) the most general
  and powerful formulation of the guarantees offered by a capability machine.
  The theorem also expresses the specific guarantees offered by the machine for
  local capabilities. It is very general and not tied to a specific ABI or a
  specific way of using the system's capabilities.
\item The above theorem is formulated in terms of a unary step-indexed
  logical relation with recursively defined Kripke worlds. It is
  inspired by one used in \cite{Devriese:2016ObjCap} for
  capability-safety in a JavaScript-like lambda calculus, but has two
  novel technical characteristics. The natural CPS style of assembly
  code with return pointers as continuations is modeled using a
  \emph{single} orthogonal closure, rather than the biorthogonal
  closure that has been previously used for lambda calculi. Perhaps
  more surprisingly, we show that the special nature of local
  capabilities can be elegantly expressed using a variant of an idea
  that has been used by Dreyer et. al.~\cite{Dreyer:jfp12} to express
  well-bracketedness guarantees in a lambda calculus: Kripke worlds
  with both a weak and a strong future world relation. Interestingly,
  the two future world relations generally work very similarly to
  Dreyer et al.'s, except that they turn up in quite a different place
  in the logical relation.
\item We demonstrate the power of our results by applying them to challenging
  examples, specifically constructed to demonstrate local state encapsulation
  and control-flow integrity guarantees in the presence of cross-component
  function pointers. The examples demonstrate both the power of our formulation
  of capability-safety and our stack management discipline.
\end{itemize}

>>>>

In this paper, we make progress towards the goal of efficient and provably
watertight enforcement of encapsulation and control-flow integrity on capability
machines.  Specifically, we make the following contributions:
\begin{itemize}
\item We formalise a simple but representative capability machine featuring
  local capabilities much like CHERI's, and define its operational semantics.
\item We define a novel stack management discipline for enforcing control-flow
  integrity and encapsulation of local state on the stack. The discipline relies
  solely on local capabilities (used to represent stack and return pointers) and
  does not require OS support in the form of a trusted shadow stack or special
  call and return instructions. It also does not require separate per-component
  stacks and instead works with a single, fairly standard, contiguous stack. It
  supports higher-order cross-component calls (e.g., (potentially)
  cross-component function pointers) and can be efficient assuming
  only \lars{can we say only here ?}
  one additional piece of processor support: an instruction for efficiently clearing
  a range of memory.
\item We provide a theorem that is (by far) the most general
  and powerful formulation of the guarantees offered by a capability machine.
  The theorem also expresses the specific guarantees offered by the machine for
  local capabilities. It is very general and not tied to a specific ABI or a
  specific way of using the system's capabilities.
\item The above theorem is formulated in terms of a unary step-indexed
  logical relation with recursively defined Kripke worlds. It is
  inspired by one used in \cite{Devriese:2016ObjCap} for
  capability-safety in a JavaScript-like lambda calculus, but has two
  novel technical characteristics. The natural CPS style of assembly
  code with return pointers as continuations is modeled using a
  \emph{single} orthogonal closure, rather than the biorthogonal
  closure that has been previously used for lambda calculi. Perhaps
  more surprisingly, we show that the special nature of local
  capabilities can be elegantly expressed using a variant of an idea
  that has been used by Dreyer et. al.~\cite{Dreyer:jfp12} to express
  well-bracketedness guarantees in a lambda calculus: Kripke worlds
  with both a public and a private future world relation. Interestingly,
  the two future world relations generally work very similarly to
  Dreyer et al.'s, except that they turn up in quite a different place
  in the logical relation.
\item We demonstrate the power of our results by applying them to challenging
  examples, specifically constructed to demonstrate local state encapsulation
  and control-flow integrity guarantees in the presence of cross-component
  function pointers. The examples demonstrate both the power of our formulation
  of capability-safety and our stack management discipline.
\end{itemize}
